---
title: "Hardware Detection and Validation"
author: "Stefano Cacciatore"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hardware Detection and Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`gpuinfo` detects CPU and GPU hardware and reports whether local compute
backends appear usable. Detection is framework-independent: a positive result
does not imply that torch, TensorFlow, or another package was compiled with
that backend.

## Quick check on the current computer

```{r quick-check, eval=FALSE}
library(gpuinfo)

has_cpu()
has_gpu()
has_cuda()
has_metal()
has_rocm()
has_opencl()

gpu_info()
backend_info()
gpu_sitrep()
```

All probes are defensive. Missing hardware, commands, drivers, or libraries
produce `FALSE`, empty results, or `NA` fields instead of an installation or
runtime error. Detailed backend information distinguishes `available`,
`unavailable`, and `unknown`; convenience predicates only return `TRUE` for a
confirmed usable backend.

## Evidence shipped with this release

Real-device results are stored with the package and returned by
`validation_info()`. The table is intentionally narrower than the implemented
probe set.

```{r validation-evidence}
library(gpuinfo)
validation_info()
```

For version 0.1.0, real GPU validation covers:

- Apple M3 with Metal on macOS;
- NVIDIA Tesla T4 with CUDA on Linux; and
- NVIDIA Tesla T4 OpenCL enumeration and CUDA/OpenCL physical-device
  deduplication on Linux.

Linux ARM64 validates the CPU path and native package compilation. Hosted
Linux, Windows, Intel macOS, and Apple Silicon macOS jobs exercise package
installation, tests, and no-GPU behavior. A hosted Windows CPU runner is not
evidence of Windows NVIDIA support.

AMD ROCm, Intel GPU OpenCL, and Windows NVIDIA CUDA/OpenCL remain unvalidated
for version 0.1.0. Their probes are available for early testing but are not
represented as validated hardware support.

## Reproducing a real-hardware run

Use a clean public commit so the tested source is recoverable. Replace
`BACKEND` with `cuda`, `metal`, `rocm`, or `opencl`.

```{sh hardware-validation, eval=FALSE}
git clone https://github.com/tkcaccia/gpuinfo.git
cd gpuinfo
git rev-parse HEAD
R CMD INSTALL .
Rscript tools/validate-hardware.R BACKEND | tee validation-output.txt
```

A successful run ends with `VALIDATION PASSED`. Preserve the full output and
record the package commit, operating system, architecture, GPU model, driver,
runtime, memory, provider, and job identifier. Cloud instances must be
terminated after collecting the evidence; merely disconnecting from or
stopping a billable GPU machine may leave charges running.

## What counts as validation

The project uses three distinct levels of testing:

1. Parser fixtures verify that representative command output is interpreted
   correctly. They do not prove hardware support.
2. CPU-only continuous integration verifies portability, installation, and
   safe behavior when accelerators are absent. It does not validate a GPU.
3. A real-hardware run must enumerate the device and expected backend and must
   complete `tools/validate-hardware.R` successfully. Only these runs are added
   to `validation_info()`.

This separation keeps release claims auditable and allows later package
versions to promote AMD, Intel, or Windows GPU support only after the relevant
hardware run passes.
